---
title: "numform"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  md_document:
    toc: true      
---

```{r, echo=FALSE}
pacman::p_load(numform, xtable, wakefield, ggplot2, knitr)
rinline <- function(code) { 
    sprintf('<code class="r">``` `r %s` ```</code>', code) 
} 
desc <- suppressWarnings(readLines("DESCRIPTION"))
regex <- "(^Version:\\s+)(\\d+\\.\\d+\\.\\d+)"
loc <- grep(regex, desc)
ver <- gsub(regex, "\\2", desc[loc])
verbadge <- sprintf('<a href="https://img.shields.io/badge/Version-%s-orange.svg"><img src="https://img.shields.io/badge/Version-%s-orange.svg" alt="Version"/></a></p>', ver, ver)
````

```{r, echo=FALSE, message=FALSE, warning=FALSE}
knit_hooks$set(htmlcap = function(before, options, envir) {
  if(!before) {
    paste('<p class="caption"><b><em>',options$htmlcap,"</em></b></p>",sep="")
    }
    })
knitr::opts_knit$set(self.contained = TRUE, cache = FALSE)
knitr::opts_chunk$set(fig.path = "inst/figure/")
```

[![Build Status](https://travis-ci.org/trinker/numform.svg?branch=master)](https://travis-ci.org/trinker/numform)
[![Coverage Status](https://coveralls.io/repos/trinker/numform/badge.svg?branch=master)](https://coveralls.io/r/trinker/numform?branch=master) 
`r verbadge`

**numform** contains tools to assist in the formatting of numbers for publication.  Tools include the removal of leading zeros, standardization of number of digits, and a p-value formatter.

# Installation

To download the development version of **numform**:

Download the [zip ball](https://github.com/trinker/numform/zipball/master) or [tar ball](https://github.com/trinker/numform/tarball/master), decompress and run `R CMD INSTALL` on it, or use the **pacman** package to install the development version:

```r
if (!require("pacman")) install.packages("pacman")
pacman::p_load_current_gh("trinker/numform")
pacman::p_load(tidyverse)
```

# Contact

You are welcome to:
* submit suggestions and bug-reports at: <https://github.com/trinker/numform/issues>
* send a pull request on: <https://github.com/trinker/numform>
* compose a friendly e-mail to: <tyler.rinker@gmail.com>

# Available Functions

Below is a table of available **numform** functions.  Note that `f_` is read as "format" whereas `fv_` is read as "format vector".  The former formats individual values in the vector while the latter uses the vector to compute a calculation on each of the values and then formats them.  Additionally, all **numform** non-methods functions have a functional return version that is prefixed with an additional `f`.  For example, `f_num` has `ff_num` which has the same arguments but returns a function instead.  This is useful for passing in to **ggplot2** `scale_x/y_type` functions (see [Plotting](#plotting) for usage).


```{r, results='asis', echo=FALSE, comment=NA, warning=FALSE, htmlcap="Available Formatting Functions"}
p_funs(numform) %>%
    {grep("^ff", ., invert = TRUE, value = TRUE)} %>%
    wakefield:::variables_as_matrix() %>%
    xtable::xtable() %>%
    print(type = 'html', include.colnames = FALSE, include.rownames = FALSE,
        html.table.attributes = '')
```


# Demonstration

## Load Packages

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load_gh("trinker/numform")
pacman::p_load(dplyr)
```

## Numbers

```{r, warn=FALSE}
f_num(c(0.0, 0, .2, -00.02, 1.122222, pi, "A"))
```

## Abbreviated Numbers

```{r, warn=FALSE}
f_thous(1234)
f_thous(12345)
f_thous(123456)
f_mills(1234567)
f_mills(12345678)
f_mills(123456789)
f_bills(1234567891)
f_bills(12345678912)
f_bills(123456789123)
```

## Commas

```{r, warn=FALSE}
f_comma(c(1234.12345, 1234567890, .000034034, 123000000000, -1234567))
```

## Percents

```{r, warn=FALSE}
f_percent(c(30, 33.45, .1), 1)
f_percent(c(0.0, 0, .2, -00.02, 1.122222, pi))
f_prop2percent(c(.30, 1, 1.01, .33, .222, .01))
```

## Dollars

```{r, warn=FALSE}
f_dollar(c(0, 30, 33.45, .1))
f_dollar(c(0.0, 0, .2, -00.02, 1122222, pi)) %>% 
    f_comma()
```

## Tables

```{r}
pacman::p_load(dplyr, pander)

set.seed(10)
dat <- data_frame(
    Team = rep(c("West Coast", "East Coast"), each = 4),
    Year = rep(2012:2015, 2),
    YearStart = round(rnorm(8, 2e6, 1e6) + sample(1:10/100, 8, TRUE), 2),
    Won = round(rnorm(8, 4e5, 2e5) + sample(1:10/100, 8, TRUE), 2),
    Lost = round(rnorm(8, 4.4e5, 2e5) + sample(1:10/100, 8, TRUE), 2),
    WinLossRate = Won/Lost,
    PropWon = Won/YearStart,
    PropLost = Lost/YearStart
)

mills <- ff_mills(relative = 1, prefix = '$')
percents <- ff_prop2percent(digits = 0)

dat %>%
    group_by(Team) %>%
    mutate(ChangeWinLoss = fv_percent_diff(WinLossRate, 0)) %>%
    ungroup() %>%
    mutate_at(vars(YearStart:Lost), funs(mills)) %>%
    mutate_at(vars(PropWon, PropLost), funs(percents)) %>%
    mutate(
        Team = fv_runs(Team),
        WinLossRate = f_num(WinLossRate, 1, prefix = '$', pad.char = ' ')
    ) %>%
    pander::pander(split.tables = Inf, justify = c('left', rep('right', ncol(dat))))
```


## Plotting

```r
library(tidyverse)
```

```{r}
data_frame(
    revenue = rnorm(10000, 500000, 50000),
    date = sample(seq(as.Date('1999/01/01'), as.Date('2000/01/01'), by="day"), 10000, TRUE),
    site = sample(paste("Site", 1:5), 10000, TRUE)
) %>%
    mutate(
        dollar = f_dollar(revenue, digits = -3),
        thous = f_thous(revenue),
        thous_dollars = f_thous(revenue, prefix = '$'),
        abb_month = f_month(date),
        abb_week = factor(f_weekday(date, distinct = TRUE), levels = c('Su', 'M', 'T', 'W', 'Th', 'F', 'S'))
    ) %T>%
    print() %>%
    ggplot(aes(abb_week, revenue)) +
        geom_jitter(width = .2, height = 0, alpha = .2) +
        scale_y_continuous(label = ff_thous(prefix = '$'))+
        facet_wrap(~site) +
        theme_bw()
```


## Modeling

We can see its use in actual model reporting as well:


```{r}
mod1 <- t.test(1:10, y = c(7:20))

sprintf(
    "t = %s (%s)",
    f_num(mod1$statistic),
    f_pval(mod1$p.value)
)
```

```{r}
mod2 <- t.test(1:10, y = c(7:20, 200))

sprintf(
    "t = %s (%s)",
    f_num(mod2$statistic, 2),
    f_pval(mod2$p.value, digits = 2)
)
```

We can build a function to report model statistics:

```{r}
report <- function(mod, stat = NULL, digits = c(0, 2, 2)) {
    
    stat <- if (is.null(stat)) stat <- names(mod[["statistic"]])
    sprintf(
        "%s(%s) = %s, %s", 
        gsub('X-squared', '&Chi;<sup>2</sup>', stat),
        paste(f_num(mod[["parameter"]], digits[1]), collapse = ", "),
        f_num(mod[["statistic"]], digits[2]),
        f_pval(mod[["p.value"]], digits = digits[3])
    )

}

report(mod1)
report(oneway.test(count ~ spray, InsectSprays))
report(chisq.test(matrix(c(12, 5, 7, 7), ncol = 2)))
```

This enables in-text usage as well.  First set up the models in a code chunk:

```{r}
mymod <- oneway.test(count ~ spray, InsectSprays)
mymod2 <- chisq.test(matrix(c(12, 5, 7, 7), ncol = 2))
```

And then use `r rinline("report(mymod)")` resulting in a report that looks like this: `r report(mymod)`.  For &Chi;<sup>2</sup> using proper HTML leads to `r report(mymod2)`.
